\documentclass[11pt,a4paper]{ltxdoc}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{array}
\usepackage{listings}
\usepackage{fancyvrb}
\usepackage{url}

% Define custom blue color
\definecolor{customblue}{HTML}{4169E1}

% Configure hyperref with custom colors
\usepackage[
  colorlinks=true,
  linkcolor=customblue,
  urlcolor=customblue,
  citecolor=customblue,
  filecolor=customblue,
  bookmarksnumbered=true,
  bookmarksopen=true
]{hyperref}

% Package being documented
\usepackage[backend=biber]{biblatex}
\addbibresource{../tests/test-asa.bib}

\lstset{
  basicstyle=\ttfamily\small,
  commentstyle=\color{gray},
  keywordstyle=\color{blue},
  stringstyle=\color{red},
  showstringspaces=false,
  breaklines=true,
  frame=single,
  backgroundcolor=\color{gray!10}
}

\title{The \textsf{biblatex-asa} Package\\
       \Large American Sociological Association Style for \textsf{biblatex}}
\author{%
  \begin{tabular}{@{}c@{\hspace{3cm}}c@{}}
    Yuqi Liang & Version 1.0.0 \\
    \texttt{dawson1900@live.com} & July 4, 2025
  \end{tabular}%
}
\date{}

\begin{document}

\maketitle

\begin{abstract}
The \textsf{biblatex-asa} package provides a complete implementation of the American Sociological Association (ASA) citation and bibliography style for LaTeX using the \textsf{biblatex} package. This implementation follows the format requirements of the \emph{American Sociological Review} and the ASA Style Guide.
\end{abstract}

\tableofcontents

\section{Introduction}

The \textsf{biblatex-asa} package provides citation and bibliography formatting according to the American Sociological Association (ASA) style guide. This package is particularly useful for students writing sociology papers or theses, researchers submitting to sociology journals, and anyone requiring ASA-style citations and references.

\subsection{Features}

\begin{itemize}
  \item Complete ASA citation and bibliography formatting
  \item Support for all major entry types (articles, books, datasets, etc.)
  \item Proper handling of multiple authors and publication dates
  \item Correct page number formatting for different entry types
  \item Full integration with \textsf{biblatex}
  \item Unicode support via \texttt{biber} backend
\end{itemize}

\subsection{Requirements}

You will need:

\begin{itemize}
  \item \textbf{LaTeX distribution:} TeX Live 2020 or later, MiKTeX 2020 or later
  \item \textbf{biblatex:} Version 3.14 or later
  \item \textbf{biber:} Version 2.14 or later (recommended backend)
  \item \textbf{etoolbox:} Version 2.5 or later (automatically loaded by biblatex)
\end{itemize}

The package is built on top of biblatex's \texttt{standard} and \texttt{authoryear} styles.

\section{Installation}

\subsection{Automatic Installation}

Install through your TeX distribution's package manager:

\begin{itemize}
  \item \textbf{TeX Live}: \texttt{tlmgr install biblatex-asa}
  \item \textbf{MiKTeX}: Available through MiKTeX Package Manager
  \item \textbf{Overleaf}: Pre-installed and ready to use
\end{itemize}

\subsection{Manual Installation}

For local use, copy these files to your project directory:
\begin{itemize}
  \item \texttt{biblatex-asa.sty}
  \item \texttt{asa.bbx}
  \item \texttt{asa.cbx}
\end{itemize}

For system-wide installation, place files in your local \texttt{texmf} tree under \texttt{texmf/tex/latex/biblatex-asa/} and run \texttt{texhash} (TeX Live) or \texttt{initexmf --update-fndb} (MiKTeX).

\section{Usage}

\subsection{Basic Usage}

Load the package in your document preamble:

\begin{lstlisting}[language=TeX]
\documentclass{article}
\usepackage{biblatex-asa}
\addbibresource{references.bib}

\begin{document}
This is a citation \parencite{key2023}.
\textcite{author2022} argues that...
\printbibliography
\end{document}
\end{lstlisting}

\subsection{Direct Style Usage}

Alternatively, use the ASA style directly:

\begin{lstlisting}[language=TeX]
\usepackage[backend=biber,style=asa]{biblatex}
\addbibresource{references.bib}
\end{lstlisting}

\subsection{Package Options}

\begin{description}
  \item[\texttt{giveninits}] Use initials for given names
  \item[\texttt{nogiveninits}] Use full given names (default)
  \item[\texttt{backend=biber}] Use biber backend (recommended)
  \item[\texttt{backend=bibtex}] Use bibtex backend
\end{description}

Example:
\begin{lstlisting}[language=TeX]
\usepackage[giveninits,backend=biber]{biblatex-asa}
\end{lstlisting}

\section{Citation Commands}

The package supports standard \textsf{biblatex} citation commands:

\begin{description}
  \item[\cs{parencite\{key\}}] Parenthetical citation: (Author Year)
  \item[\cs{textcite\{key\}}] In-text citation: Author (Year)
  \item[\cs{cite\{key\}}] Bare citation: Author Year
\end{description}

Examples:
\begin{itemize}
  \item \texttt{\cs{parencite\{smith2020\}}} → (Smith 2020)
  \item \texttt{\cs{textcite\{jones2019\}} argues...} → Jones (2019) argues...
  \item \texttt{\cs{parencite[p.~42]\{doe2018\}}} → (Doe 2018, p. 42)
\end{itemize}

\section{Compilation}

Compile documents using this sequence:

\begin{enumerate}
  \item \texttt{pdflatex document.tex}
  \item \texttt{biber document}
  \item \texttt{pdflatex document.tex}
  \item \texttt{pdflatex document.tex}
\end{enumerate}

For Overleaf users, the compilation sequence is handled automatically.

\section{ASA Format Implementation}

\subsection{Citation Format}

\begin{itemize}
  \item Single author: (Smith 2020)
  \item Two authors: (Smith and Jones 2020)
  \item Three or more authors: (Smith et al. 2020)
  \item Multiple citations: (Smith 2020; Jones 2021)
\end{itemize}

\subsection{Bibliography Format}

\begin{itemize}
  \item Authors: First author inverted (Last, First), subsequent authors normal order
  \item All authors listed (no "et al." truncation in bibliography)
  \item Articles: Journal name italicized, volume(issue):pages format
  \item Books: Publisher and location information
  \item Book chapters: "Pp. X–Y in..." formatting
  \item Proper handling of publication dates, DOIs, and URLs
\end{itemize}

\section{Package Details}

The detailed implementation and examples for this package are contained in the source files and example documents distributed with this package:

\begin{description}
  \item[\texttt{biblatex-asa.sty}] The main package file that provides a convenient wrapper interface. This file loads the ASA style and sets up appropriate defaults for American Sociological Association formatting.

  \item[\texttt{asa.bbx}] The bibliography style file that defines how references appear in the bibliography section. This file contains the core formatting logic for ASA-style reference lists, including author ordering, journal formatting, and special handling for different entry types.

  \item[\texttt{asa.cbx}] The citation style file that controls how citations appear in the text. This implements ASA parenthetical citation format, including proper handling of multiple authors, page numbers, and year disambiguation.

  \item[\texttt{examples/basic/example.tex}] A basic example document demonstrating standard usage of the ASA style. This example shows fundamental citation commands and bibliography generation using direct biblatex style loading.

  \item[\texttt{examples/with-package/example-with-package.tex}] An example using the convenience package wrapper. This demonstrates the recommended approach for most users and includes various citation scenarios and entry types.

  \item[\texttt{tests/test-asa.tex}] A comprehensive test document that exercises all supported entry types and citation formats. This document serves as both a validation tool and a reference for advanced formatting scenarios.

  \item[\texttt{tests/test-asa.bib}] The bibliography database containing test entries for all major bibliographic entry types. These entries demonstrate proper field usage and formatting expectations for ASA style.
\end{description}

Users new to the package should start with the \texttt{with-package} example, which provides a complete working example with detailed comments. Advanced users and package developers may find \texttt{tests/test-asa.tex} useful for understanding edge cases and comprehensive formatting options.

\section{Supported Entry Types}

The package supports all standard \textsf{biblatex} entry types with ASA formatting:

\begin{itemize}
  \item \texttt{@article} - Journal articles
  \item \texttt{@book} - Books
  \item \texttt{@incollection} - Book chapters
  \item \texttt{@inproceedings} - Conference proceedings
  \item \texttt{@misc} - Miscellaneous sources
  \item \texttt{@online} - Online sources
  \item \texttt{@report} - Reports
  \item \texttt{@thesis} - Theses and dissertations
\end{itemize}

\section{Troubleshooting}

\begin{description}
  \item[References not showing] Ensure you run \texttt{biber} (not \texttt{bibtex}).
  \item[File not found errors] Check package installation or ensure files are in project directory.
  \item[Encoding problems] Save all files as UTF-8.
\end{description}

\section{License and Contact}

This work is distributed under the LaTeX Project Public License (LPPL), version 1.3c or later.

\begin{description}
  \item[Author:] Yuqi Liang, University of Oxford
  \item[Email:] \texttt{dawson1900@live.com}
  \item[GitHub:] \url{https://github.com/yuqi-liang-qiqi/biblatex-asa}
\end{description}

\end{document} 